Designing Types for R, Empirically

### 2020-05-15 evaluation/R/corpus-analysis.Rmd
Updated 5

```{r}
pkgs <- read_csv(
  CRAN_PACKAGES_DETAILS,
  col_types=cols(
    package = col_character(),
    installed = col_logical(),
    loadable = col_logical(),
    version = col_character(),
    title = col_character(),
    size = col_double(),
    package_native_code = col_double(),
    package_r_code = col_double(),
    package_code = col_double(),
    runnable_code = col_double(),
    runnable_files = col_double(),
    runnable_code_examples = col_double(),
    runnable_code_vignettes = col_double(),
    runnable_code_tests = col_double(),
    runnable_files_examples = col_double(),
    runnable_files_vignettes = col_double(),
    runnable_files_tests = col_double(),
    coverage_expr = col_double(),
    coverage_line = col_double(),
    runnable_success = col_double(),
    runnable_error = col_double(),
    runnable_time = col_double(),
    runnable_rate = col_double(),
    funs_private = col_double(),
    funs_public = col_double(),
    funs = col_double(),
    revdeps = col_double(),
    funs_with_assert = col_double(),
    asserts = col_double(),
    uses_roxygen = col_logical()
  )
)
```

@@

    r(paste(prefix, "with coverage"), w_coverage),
    r(paste(prefix, "mean expr coverage"), percent(mean(w_coverage))),
    r(paste(prefix, "median expr coverage"), percent(median(w_coverage))),

@@

    r(paste(prefix, "functions"), sum(df$funs, na.rm=T)),
    r(paste(prefix, "public functions"), sum(df$funs_public, na.rm=T)),
    r(paste(prefix, "private functions"), sum(df$funs_private, na.rm=T)),

    r(paste(prefix, "w revdes"), w_revdeps),
    r(paste(prefix, "mean revdes"), mean(w_revdeps)),
    r(paste(prefix, "median revdes"), median(w_revdeps)),
    r(paste(prefix, "at least min revdep"), filter(w_revdeps, revdeps >= params$min_revdeps)),

>>>>

```{r}
pkgs <- read_csv(CRAN_PACKAGES_DETAILS)
```

@@

    r(paste(prefix, "with coverage"), w_coverage),
    r(paste(prefix, "mean expr coverage"), percent(mean(w_coverage$coverage_expr))),
    r(paste(prefix, "median expr coverage"), percent(median(w_coverage$coverage_expr))),
    r(paste(prefix, "covered functions ratio"), ratio(sum(w_coverage$funs_with_coverage),sum(w_coverage$funs))),

@@

    r(paste(prefix, "functions"), sum(df$funs, na.rm=T)),
    r(paste(prefix, "public functions"), sum(df$funs_public, na.rm=T)),
    r(paste(prefix, "private functions"), sum(df$funs_private, na.rm=T)),
    r(paste(prefix, "S three functions"), sum(df$funs_s3, na.rm=T)),
    r(paste(prefix, "S three classes"), sum(df$s3_classes, na.rm=T)),

    r(paste(prefix, "w revdes"), w_revdeps),
    r(paste(prefix, "revdes"), sum(w_revdeps$revdeps)),
    r(paste(prefix, "mean revdes"), mean(w_revdeps$revdeps)),
    r(paste(prefix, "median revdes"), median(w_revdeps$revdeps)),
    r(paste(prefix, "at least min revdep"), filter(w_revdeps, revdeps >= params$min_revdeps)),
    r(paste(prefix, "revdep runnable code"), sum(w_revdeps$revdeps_runnable_code, na.rm=T)),
    r(paste(prefix, "revdep mean code coverage"), percent(mean(w_revdeps$revdep_coverage_expr)/100)),
    r(paste(prefix, "revdep median code coverage"), percent(median(w_revdeps$revdep_coverage_expr)/100)),

### 2020-05-15 paper/main.tex
Updated 5

For this paper we have selected \CorpusLoadable packages consisting of
\CorpusRCodeRnd lines of R code and \CorpusNativeCodeRnd lines of native
code (C/C++/Fortran). Figure~\ref{fig:corpus} shows all the packages, the
size of the dots reflects the size of the project in terms of both R and
native lines of code\footnote{ Lines of source code reported excludes
  comments and blank lines, counted by \emph{cloc}, \cf
  \url{https://github.com/AlDanial/cloc}}, the x-axis indicates the
expression code coverage in percent and the y-axis gives the number of
reverse dependencies in log scale. Dotted lines indicate means.  Packages
with over \PackageSizeOutierRnd lines of code are annotated.

\begin{figure}[!h]  \centering
  \includegraphics[width=.98\linewidth]{plots/corpus.pdf}
  \caption{Corpus overview}\label{fig:corpus}
\end{figure}

These packages come from the Comprehensive R Achive Network
(CRAN\footnote{\url{http://cran.r-project.org}}), the largest repository of
R code with over \AllCranRnd packages\footnote{CRAN receives about 6 new
  package submissions a day~\cite{Ligges2017}} containing over \AllRCodeRnd
and \AllNativeCodeRnd R and native code respectively. Unlike other open code
repositories like GitHub, CRAN is a curated repository. Each submitted
package must abide to a number of well-formedness rules that are
automatically checked asserting certain quality. Most relevant for this work
is that all of the runnable code (including code snippets from examples and
vignettes) is tested and only a successfully running package is admitted in
the archive.

We have downloaded and installed all available CRAN packages. Out of the
\AllCranRnd packages, we managed to install \AllLoadableRnd. The main reason
for this is that \rdt is based on R 3.5.0 and some of the packages are no
longer compatible with it. Some packages also require extra native
dependencies which were not present on our servers.  We have two criteria
for including a package into the corpus:
\begin{inparaenum}[(1)]
\item have runnable code that covers significant part of the package source
  code from which type signatures could be inferred, and
\item have some reverse dependencies that allows us to evaluate the inferred
  types on runnable code from these dependent packages.
\end{inparaenum}
The concrete thresholds were set to at least \ThresholdCodeCoverage of
expression coverage and at minimum \ThresholdRevdeps reverse dependencies.

The code coverage was computed for each package using
\covr\footnote{\url{https://github.com/r-lib/covr}}, the R code coverage R
tool.  The reverse package dependencies were extracted from the package
metadata using builtin function.

%
%
%
\section{Left overs}

>>>>

For this paper we have selected \CorpusLoadable packages consisting of
\CorpusRCodeRnd lines of R code and \CorpusNativeCodeRnd lines of native code
(C/C++/Fortran). Figure~\ref{fig:corpus} shows these packages, the size of the
dots reflects the size of the project in terms of lines of code including both R
and native code\footnote{ Lines of source code reported excludes comments and
  blank lines, counted by \emph{cloc}, \cf
  \url{https://github.com/AlDanial/cloc}}, the x-axis indicates the expression
code coverage in percents and the y-axis gives the number of reverse
dependencies in log scale. Dotted lines indicate means. Packages with over
\PackageSizeOutierRnd lines of code are annotated.

\begin{figure}[!h]  \centering
  \includegraphics[width=.8\linewidth]{plots/corpus.pdf}
  \caption{Corpus overview}\label{fig:corpus}
\end{figure}

These packages come from the Comprehensive R Achive Network
(CRAN\footnote{\url{http://cran.r-project.org}}), the largest repository of R
code with over \AllCranRnd packages\footnote{CRAN receives about 6 new package
  submissions a day~\cite{Ligges2017}} containing over \AllRCodeRnd and
\AllNativeCodeRnd R and native code respectively. Unlike other open code
repositories such as GitHub, CRAN is a curated repository. Each submitted
package must abide to a number of well-formedness rules that are automatically
checked asserting certain quality. Most relevant for this work is that all of
the runnable code (including code snippets from examples and vignettes) is
tested and only a successfully running package is admitted in the archive.

We have downloaded and installed all available CRAN packages. Out of the
\AllCranRnd packages, we managed to install \AllLoadableRnd. The main reason for
this is that \rdt is based on R 3.5.0 and some of the packages are no longer
compatible with this version. Some packages also require extra native
dependencies which were not present on our servers. We defined two criteria for
including a package into the corpus:
\begin{inparaenum}[(1)]
  \item have a runnable code that covers a significant part of the package
  source code from which type signatures could be inferred, and
  \item have some reverse dependencies that will allows us to evaluate the
  inferred types using the runnable code from these dependencies.
\end{inparaenum}
The concrete thresholds used were: at least \ThresholdCodeCoverage of expression
coverage and at minimum \ThresholdRevdeps reverse dependencies. The code
coverage was computed for each package using
\covr\footnote{\url{https://github.com/r-lib/covr}}, the R code coverage R tool.
The reverse package dependencies were extracted from the package metadata using
builtin function.

The \CorpusLoadable selected packages contain \CorpusRunnableCodeRnd lines of
runnable code in examples (\CorpusExamplesCodeRnd), tests (\CorpusTestsCodeRnd)
and vignettes (\CorpusVignettesCodeRnd). Running this code results on average in
\CorpusMeanExprCoverage package code coverage (CRAN average is
\AllMeanExprCoverage). Together, there is \CorpusRevdesRnd (on average
\CorpusMeanRevdes, median \CorpusMedianRevdes; CRAN average is \AllMeanRevdes,
median \AllMedianRevdes) reverse dependencies with \CorpusRevdepRunnableCodeRnd
runnable lines of code resulting in \CorpusRevdepMeanCodeCoverage coverage (on
average) of the corpus packages.

Together there are \CorpusFunctionsRnd defined R functions
(\CorpusPublicFunctionsRnd are from packages' public API).
\CorpusSThreeFunctionsRnd are S3 functions. Packages in the corpus define
\CorpusSThreeClasses S3 classes.

%
%
\section{Left overs}
